<P>

<BR>

<CENTER>

<FONT size=5.5, color="Red">
**M3 Project Report** </FONT>

<FONT size=5.5, color="#800080">
<BR>**Probability Theory and Introductory Statistics** </FONT>


<P>

<FONT size=5.5, color="#008080"> ALY 6010 - 71579 </FONT>


<P>

<BR>
<FONT size=5, color="Grey"> 
<BR>
**Maheswar Raju Narasaiah**

<FONT size=5, color="Black"> 
Prof: Dr. Dee Chiluiza 

Date:  `r format(Sys.time(), '%d %B, %Y')`

</CENTER>

<P>
<BR> <B>
<FONT SIZE = 5, COLOR ="#f70d1a">
1. INTRODUCTION
</FONT>
</BR></B>
   
   <FONT SIZE = 4.75, COLOR ="#f94c56">
   **A. Confidence Interval, it's Importance and Application**
   </FONT>
   <P>
<FONT SIZE = 4.5>


A confidence interval (CI) is a set of numbers that includes the value of a previously unknown population parameter. These intervals provide a plausible domain for the parameter based on the qualities of the sample data. Confidence intervals are formed from sample statistics and produced using a predetermined confidence level.

Because it is often hard to measure entire populations, population parameters are highly variable. A sample can be used to estimate these parameters. Due to random sampling error, however, the estimates rarely exactly equal the parameter.

The confidence level is the long-run probability that a set of confidence intervals will encompass the real value of a population parameter.

-- ***Importance of Confidence Interval***

The majority of what we do in statistics includes using a sample to answer a research question. A confidence interval expresses the degree of uncertainty or mistake associated with an estimated statistic. It should be noted that the interval can only reflect the uncertainty introduced by taking a sample from the underlying population (that is, sampling variability). A confidence interval cannot infer non-sampling concerns such as bias, measure accuracy, or generalization of the result.

Generally, the confidence interval takes the form:

<P>
  *C.I. = statistic ± multiplier × standard error of the statistic*

</P>

The "statistic" in this calculation is the estimate from our sample; it is our best guess at the true value in the population. This statistic could be a mean, a percentage, a difference in means, a regression coefficient, a relative risk, a hazard ratio, or something else. The "multiplier" is a value from a specific theoretical distribution (for example, the normal distribution or the t distribution); the suitable distribution is determined by the type of statistic. The value chosen from that distribution expresses the "confidence" that the unknown parameter will fall inside the confidence interval. The variability due to sampling is represented by the "standard error of the statistic."

Different random samples collected from the same population will almost certainly provide somewhat different intervals. If you take a large number of random samples and compute a confidence interval for each one, a percentage of them will contain the parameter.


Fortunately, inferential statistics methods may analyze a sample while compensating for the uncertainty 
that comes with collecting samples. Confidence intervals encompass the point estimate with a margin 
of uncertainty to assist us in comprehending how wrong the estimate could be.

-- ***Applications of Confidence Interval in News Industry***

When poll results are revealed, confidence intervals are frequently shown on the news. This is an example from the October 2018 issue of the CNN Press:

According to the most recent ABC News-Washington Post survey, 56 percent of respondents supported Trump, while 39 percent supported Biden. The ABC News-Washington Post telephone poll of 1,014 individuals was conducted between March 8 and 10, with a margin of error of 3.5 percentage points. (Emphasis added.)

Even though it is not indicated, the margin of error presented here was most likely the 95 percent confidence interval. Simply put, there is a 95 percent chance that between 35.5 percent and 42.5 percent of voters will vote for Biden (39 percent plus or minus 3.5 percent). In contrast, there is a 5% possibility that less than 35.5 percent of voters or more than 42.5 percent of voters will support Biden.

The precise statistical definition of the 95 percent confidence interval is that if the telephone poll were conducted 100 times, 95 percent of respondents favoring Biden would be within the calculated confidence intervals and five percent of respondents favoring Biden would be either higher or lower than the confidence interval range.

Instead of 95 percent confidence intervals, confidence intervals depending on alternative degrees of significance, such as 90 percent or 99 percent, might be used. The statistical word for how willing you are to be wrong is the level of significance. You have a 5% probability of being wrong with a 95% confidence interval. 
You have a 10% chance of being mistaken with a 90% confidence interval. A 95% confidence interval would be wider than a 95% confidence interval (for example, plus or minus 4.5 percent instead of 3.5 percent). A confidence interval of 90% would be narrower (plus or minus 2.5 percent, for example).

</FONT>

<FONT SIZE = 4.75, COLOR ="#f94c56">
   **B. Hypothesis Testing, it's Importance and Application**
   </FONT>
   <P>
<FONT SIZE = 4.5>

Hypothesis testing determines whether a specific assumption holds true for the entire population. It is a statistical instrument. It evaluates sample data from the entire population to determine the validity of conclusion.

The hypothesis idea is based on the likelihood of an event occurring. It determines whether or not the primary hypothesis results are correct. It is commonly used in biological research, criminal trials, marketing, and manufacturing.

-- ***Importance of Hypothesis Testing***

The most important advantage of hypothesis testing is that it allows you to assess the strength of your claim or assumption before including it into your data collection. Furthermore, hypothesis testing is the sole valid approach for determining if something "is or is not." Other advantages include:

- Hypothesis testing gives a dependable foundation for making data conclusions for your target population.
- It aids the researcher in extrapolating data from the sample to the larger population.
- The researcher can use hypothesis testing to examine whether the data from the sample is statistically significant.
- In any systematic inquiry, hypothesis testing is one of the most crucial stages for determining the validity and dependability of results.
- It is beneficial to provide connections between the underlying theory and specific research concerns.


-- ***Applications of Hypothesis Testing***

In biology, hypothesis testing are frequently used to assess whether a new treatment, fertilizer, pesticide, chemical, etc. produces greater growth, stamina, immunity, and so on in plants or animals.

For example, imagine a biologist believes that a specific fertilizer will allow plants to grow 20 inches faster in a month than they generally do. For one month, she applies the fertilizer to each of the plants in her experiment.

She then performs a hypothesis test using the following hypotheses:

H0: μ = 20 inches (the fertilizer will have no effect on the mean plant growth)
HA: μ > 20 inches (the fertilizer will cause mean plant growth to increase)
If the p-value of the test is less than a certain level of significance (e.g. =.05), she can reject the null hypothesis and infer that the fertilizer increases plant growth.

</FONT>


<P>
<BR> <B>
  <FONT SIZE = 5, COLOR ="#f70d1a">
  1. ANALYSIS 
  </FONT>
</BR></B>

<P>

<FONT SIZE = 4.5>
In this below section, we will perform various statistical analysis on our Data to make inferences and address our problems.
<FONT/>

```{r message=FALSE, warning=FALSE}

# Library Used for
#####################################
library(magrittr)
library(knitr)
library(plotrix)
library(plyr)
library(dplyr)
library(readxl)
library(gridExtra)
library(CGPfunctions)
library(RColorBrewer)
library(lattice)
library(ggplot2)
library(DT)
library(kableExtra)




# Datasets used in Project"
###################################
## Sample data
SData <- read_excel("~/Desktop/Intro To Analytics - ALY 6000/ALY 6000 - Project/Data Sets/M3ProjectData.xlsx")

## Removing NA Values from each column using na.omit() function
S1 <- na.omit(SData$`Sample 1`)
S2 <- na.omit(SData$`Sample 2`)
S3 <- na.omit(SData$`Sample 3`)
S4 <- na.omit(SData$`Sample 4`)
S5 <- na.omit(SData$`Sample 5`)
S6 <- na.omit(SData$`Sample 6`)
```

***About the Data Set***
The data contains population and sample taken from population and placed into different columns


<P>
<BR> <B>
<FONT SIZE = 4.75, COLOR ="#f94c56">
2.1. Presenting First 15 Observations of Data Set
</FONT>
</BR></B>

<FONT SIZE = 4.5>
In this section, we will perform the head function to first 15 observation from our data
</FONT>

```{r message=FALSE, warning=FALSE}

# Creating a new objects for fetching first 15  records
hSData <- head(SData, 15)



# Present the table using kableExtra Package
hSData %>%
  kbl(
    caption = "Table 1 : First 15 Observation of Data Set",
    format = "html"
  ) %>%
  kable_material_dark(
    full_width = F,
    html_font = "Times New Roman"
  )
```

<BR>
<FONT SIZE = 4.5>

***Observations***

</FONT>

<P>
<BR> <B>
<FONT SIZE = 4.75, COLOR ="#f94c56">
2.2. Basic Descriptive Statistics : Population and Sample 1
</FONT>
</BR></B>


```{r message=FALSE, warning=FALSE}


# Population : Basic Descriptive Statistics
MeanPop <- mean(SData$Population)
MedianPop <- median(SData$Population)
StdPop <- sd(SData$Population)


## Vector Creation for Population
vectorPop <- c(MeanPop, MedianPop, StdPop)


# Sample 1 : Basic Descriptive Statistics
MeanS1 <- mean(S1)
MedianS1 <- median(S1)
StdS1 <- sd(S1)


## Vector Creation for Sample 1
vectorS1 <- c(MeanS1, MedianS1, StdS1)


# Matrix Creation for basic descriptive statistics
SData_Table <- matrix(c(vectorPop, vectorS1),
  nrow = 2,
  byrow = TRUE
)


## Rounding Digits in Matrix
SData_Table_Rounded <- round(SData_Table, digits = 2)

## Naming Columns and Row Names in Table
colnames(SData_Table_Rounded) <- c("Mean", "Median", "StDev") # nolint

rownames(SData_Table_Rounded) <- c("Population", "Sample 1")


# Table Creation of of Matrix Using Knitr Package
SData_Table_Rounded %>%
  knitr::kable(
    caption = "Table 2 : Descriptive Statistics of Populations and Sample 1",
    format = "html"
  ) %>%
  kable_material_dark(
    full_width = F,
    html_font = "Times New Roman"
  )
```

<BR>
<FONT SIZE = 4.5>

***Observations***

</FONT>

<P>
<BR> <B>
<FONT SIZE = 4.75, COLOR ="#f94c56">
2.3. Sample 1 : Confidence Interval for Sample Mean
</FONT>
</BR></B>


```{r message=FALSE, warning=FALSE}


# Sample 1 : Basic Descriptive Statistics
MeanS1 <- mean(S1)
StdS1 <- sd(S1)
nS1 <- 160


# Creating Object for 90% Confidence Interval
########################################
CL90S1 <- 0.90

# Calculating Alpha Value
alp90S1 <- (1 - CL90S1)

# Calculating Z value for Margin of Error
Z90LS1 <- qnorm(alp90S1 / 2)
Z90RS1 <- qnorm((alp90S1 / 2) + CL90S1)

# Calculating Margin of Error
MarErr90S1 <- Z90RS1 * (StdS1 / sqrt(nS1))

# Creating Objects for Confidence Interval: Upper and Lower Limit
CLL90S1 <- MeanS1 - MarErr90S1
CLU90S1 <- MeanS1 + MarErr90S1

# Calculation of C.I. Width for 90
wd90S1 <- CLU90S1 - CLL90S1

# Vector Creation for 90% C.I.
vectorCI90S1 <- c(Z90LS1, Z90RS1, MarErr90S1, CLU90S1, CLL90S1, wd90S1)


# Creating Object for 96% Confidence Interval
########################################
CL96S1 <- 0.96

# Calculating Alpha Value
alp96S1 <- (1 - CL96S1)

# Calculating Z value for Margin of Error
Z96LS1 <- qnorm(alp96S1 / 2)
Z96RS1 <- qnorm((alp96S1 / 2) + CL96S1)

# Calculating Margin of Error
MarErr96S1 <- Z96RS1 * (StdS1 / sqrt(nS1))

# Creating Objects for Confidence Interval: Upper and Lower Limit
CLL96S1 <- MeanS1 - MarErr96S1
CLU96S1 <- MeanS1 + MarErr96S1

# Calculation of C.I. Width for 96
wd96S1 <- CLU96S1 - CLL96S1

# Vector Creation for 96% C.I.
vectorCI96S1 <- c(Z96LS1, Z96RS1, MarErr96S1, CLU96S1, CLL96S1, wd96S1)




# Creating Object for 99% Confidence Interval
########################################
CL99S1 <- 0.99

# Calculating Alpha Value
alp99S1 <- (1 - CL99S1)

# Calculating Z value for Margin of Error
Z99LS1 <- qnorm(alp99S1 / 2)
Z99RS1 <- qnorm((alp99S1 / 2) + CL99S1)

# Calculating Margin of Error
MarErr99S1 <- Z99RS1 * (StdS1 / sqrt(nS1))

# Creating Objects for Confidence Interval: Upper and Lower Limit
CLL99S1 <- MeanS1 - MarErr99S1
CLU99S1 <- MeanS1 + MarErr99S1

# Calculation of C.I. Width for 99
wd99S1 <- CLU99S1 - CLL99S1

# Vector Creation for 99% C.I.
vectorCI99S1 <- c(Z99LS1, Z99RS1, MarErr99S1, CLU99S1, CLL99S1, wd99S1)



# Matrix Creation for compiling all the Confidence Interval Vectors
S1_Table <- matrix(c(vectorCI90S1, vectorCI96S1, vectorCI99S1), nrow = 3, byrow = TRUE)


## Rounding Digits in Matrix
S1_Table_Rounded <- round(S1_Table, digits = 2)


## Naming Columns and Row Names in Table
colnames(S1_Table_Rounded) <- c(
  "Z Left Value", "Z Right Value",
  "Marg of Error", "Upper C.L.", "Lower C.L.", "Width of C.I."
)

rownames(S1_Table_Rounded) <- c("CI 90%", "CI 96%", "CI 99%")

# Table Creation of of Matrix Using Knitr Package
S1_Table_Rounded %>%
  knitr::kable(
    caption = "Table 3: Results  of Sample 1 with
               Different Confidence Intervals",
    format = "html"
  ) %>%
  kable_classic(
    full_width = F,
    html_font = "Times New Roman"
  )
```

<BR>
<FONT SIZE = 4.5>

***Observations***

</FONT>

<P>
<BR> <B>
<FONT SIZE = 4.75, COLOR ="#f94c56">
2.4. Sample 1 : Sample Size Based On Confidence Interval
</FONT>
</BR></B>


```{r message=FALSE, warning=FALSE}

# Calculate the sample size based on known Std of Sample 1

SizeS190 <- ((Z90RS1^2) * (StdS1^2)) / (MarErr90S1^2)
SizeS196 <- ((Z96RS1^2) * (StdS1^2)) / (MarErr96S1^2)
SizeS199 <- ((Z99RS1^2) * (StdS1^2)) / (MarErr99S1^2)

# Vector Creation for size of Sample 1
SizeS1Data <- c(SizeS190, SizeS196, SizeS199)

# Calculate the sample size based on known Std of Population
SizeP90 <- ((Z90RS1^2) * (StdPop^2)) / (MarErr90S1^2)
SizeP96 <- ((Z96RS1^2) * (StdPop^2)) / (MarErr96S1^2)
SizeP99 <- ((Z99RS1^2) * (StdPop^2)) / (MarErr99S1^2)

# Vector Creation
SizePData <- c(SizeP90, SizeP96, SizeP99)



# Matrix Creation
samplesizematrix <- matrix(
  data = c(SizeS1Data, SizePData),
  nrow = 2,
  byrow = TRUE
)


## Rounding Digits in Matrix
samplesizematrix_Rounded <- round(samplesizematrix, digits = 2)

## Naming Columns and Row Names in Matrix
colnames(samplesizematrix_Rounded) <- c("CI 90% ", "CI 96%", "CI 99%")
rownames(samplesizematrix_Rounded) <- c("Sample 1", "Population")


# Table Creation of of Matrix Using Knitr Package
samplesizematrix_Rounded %>%
  knitr::kable(caption = "Table 4: Sample Size based on Known Std of Sample 1
  and Population") %>%
  kable_classic(
    full_width = F,
    html_font = "Times New Roman"
  )
```

<BR>
<FONT SIZE = 4.5>

***Observations***

</FONT>

<P>
<BR> <B>
<FONT SIZE = 4.75, COLOR ="#f94c56">
2.5. Sample 2 : Confidence Interval for Sample Mean
</FONT>
</BR></B>



```{r message=FALSE, warning=FALSE}


MeanS2 <- mean(S2)
sdS2 <- sd(S2)
nS2 <- 23

# Degree of Freedom
dfS2 <- nS2 - 1


# Calculating Confidence for 90 %
CL90S2 <- 0.90

# Calculating p Value
p90S2 <- (1 + CL90S2) / 2

# Calculating t value for Margin of Error using qt()
T90RS2 <- qt(p90S2, dfS2)
T90LS2 <- qt(p90S2 - CL90S2, dfS2)


# Calculating Margin of Error
MarErr90S2 <- T90RS2 * (sdS2 / sqrt(nS2))

# Creating Objects for Confidence Interval: Upper and Lower Limit
CLL90S2 <- MeanS2 - MarErr90S2
CLU90S2 <- MeanS2 + MarErr90S2

# Calculation of C.I. Width for 90%
wd90S2 <- CLU90S2 - CLL90S2

# Vector Creation for 90% C.I.
vectorCI90S2 <- c(T90LS2, T90RS2, MarErr90S2, CLU90S2, CLL90S2, wd90S2)



# Calculating Confidence for 96 %
CL96S2 <- 0.96

# Calculating p Value
p96S2 <- (1 + CL96S2) / 2

# Calculating t value for Margin of Error using qt()
T96RS2 <- qt(p96S2, dfS2)
T96LS2 <- qt(p96S2 - CL96S2, dfS2)


# Calculating Margin of Error
MarErr96S2 <- T96RS2 * (sdS2 / sqrt(nS2))

# Creating Objects for Confidence Interval: Upper and Lower Limit
CLL96S2 <- MeanS2 - MarErr96S2
CLU96S2 <- MeanS2 + MarErr96S2

# Calculation of C.I. Width for 96%
wd96S2 <- CLU96S2 - CLL96S2

# Vector Creation for 96% C.I.
vectorCI96S2 <- c(T96LS2, T96RS2, MarErr96S2, CLU96S2, CLL96S2, wd96S2)


# Calculating Confidence for 99 %
CL99S2 <- 0.99

# Calculating p Value
p99S2 <- (1 + CL99S2) / 2

# Calculating t value for Margin of Error using qt()
T99RS2 <- qt(p99S2, dfS2)
T99LS2 <- qt(p99S2 - CL99S2, dfS2)


# Calculating Margin of Error
MarErr99S2 <- T99RS2 * (sdS2 / sqrt(nS2))

# Creating Objects for Confidence Interval: Upper and Lower Limit
CLL99S2 <- MeanS2 - MarErr99S2
CLU99S2 <- MeanS2 + MarErr99S2

# Calculation of C.I. Width for 99%
wd99S2 <- CLU99S2 - CLL99S2

# Vector Creation for 99% C.I.
vectorCI99S2 <- c(T99LS2, T99RS2, MarErr99S2, CLU99S2, CLL99S2, wd99S2)


# Matrix Creation for compiling all the Confidence Interval Vectors
#########################
S2_Table <- matrix(c(vectorCI90S2, vectorCI96S2, vectorCI99S2),
  nrow = 3,
  byrow = TRUE
)


## Rounding Digits in Matrix
S2_Table_Rounded <- round(S2_Table, digits = 2)


## Naming Columns and Row Names in Table
colnames(S2_Table_Rounded) <- c(
  "T Left Value", "T Right Value",
  "Marg of Error", "Upper C.L.", "Lower C.L.", "Width of C.I."
)

rownames(S2_Table_Rounded) <- c("CI 90%", "CI 96%", "CI 99%")

# Table Creation of of Matrix Using Knitr Package
S2_Table_Rounded %>%
  knitr::kable(caption = "Table 5: Results of Sample 2 with
  Different Confidence Intervals", format = "html") %>%
  kable_classic(full_width = F, html_font = "Times New Roman")
```

<BR>
<FONT SIZE = 4.5>

***Observations***

</FONT>


<P>
<BR> <B>
<FONT SIZE = 4.75, COLOR ="#f94c56">
2.6. Sample 3 : Confidence Interval for Sample Proportion
</FONT>
</BR></B>


```{r message=FALSE, warning=FALSE}


# Calculation for sample proportion

nS3 <- 1500

Lower1.7S3 <- sum(S3 < 1.7)
Upper1.7S3 <- nS3 - Lower1.7S3

p_successS3 <- Lower1.7S3 / nS3
p_failureS3 <- Upper1.7S3 / nS3


# Vector Creation for Sample3
vectorS3 <- c(p_successS3, p_failureS3)


# Calculation for sample proportion

nPop <- 6556

Lower1.7Pop <- sum(SData$Population < 1.7)
Upper1.7Pop <- nPop - Lower1.7Pop



p_successPop <- Lower1.7Pop / nPop
p_failurePop <- Upper1.7Pop / nPop


# Vector Creation for Sample3
vectorPop <- c(p_successPop, p_failurePop)



# Matrix Creation for compiling all the Confidence Interval Vectors
#########################
Prop_Table <- matrix(c(vectorS3, vectorPop), nrow = 2, byrow = TRUE)


## Rounding Digits in Matrix
Prop_Table_Rounded <- round(Prop_Table, digits = 2)


## Naming Columns and Row Names in Table
colnames(Prop_Table_Rounded) <- c("Success", "Failure")

rownames(Prop_Table_Rounded) <- c("Sample 3", "Population")

# Table Creation of of Matrix Using Knitr Package
Prop_Table_Rounded %>%
  knitr::kable(
    caption = "Table 6: Proportion of Success and Failure for Sample 3 and Population",
    format = "html"
  ) %>%
  kable_classic(
    full_width = F,
    html_font = "Times New Roman"
  )

# Creating Objects
nS3 <- 1500

# To find the find the count of obseervation lower than 1.7
Lower1.7S3 <- sum(S3 < 1.7)
Upper1.7S3 <- nS3 - Lower1.7S3

# Calculating the proportion for success and failure
p_successS3 <- Lower1.7S3 / nS3
p_failureS3 <- Upper1.7S3 / nS3


# Calculating Confidence for 90 %
####################################################
CL90S3 <- 0.90

# Calculation of Z Value for 90% CI
zvalue90S3 <- qnorm((1 + CL90S3) / 2)

# Calculation of Error for 90% CI
error90S3 <- zvalue90S3 * sqrt(((p_successS3 * p_failureS3) / nS3))

# Calculation of Upper and Lower Limit for 90% CI
CLL90S3 <- p_successS3 - error90S3
CLU90S3 <- p_successS3 + error90S3

# Calculation of C.I. Width for 90%
wd90S3 <- CLU90S3 - CLL90S3

# Vector Creation for 90% C.I.
vectorCI90S3 <- c(zvalue90S3, error90S3, CLU90S3, CLL90S3, wd90S3)


# Calculating Confidence for 96 %
####################################################
CL96S3 <- 0.96

# Calculation of Z Value for 96% CI
zvalue96S3 <- qnorm((1 + CL96S3) / 2)

# Calculation of Error for 96% CI
error96S3 <- zvalue96S3 * sqrt(((p_successS3 * p_failureS3) / nS3))

# Calculation of Upper and Lower Limit for 96% CI
CLL96S3 <- p_successS3 - error96S3
CLU96S3 <- p_successS3 + error96S3

# Calculation of C.I. Width for 96%
wd96S3 <- CLU96S3 - CLL96S3

# Vector Creation for 96% C.I.
vectorCI96S3 <- c(zvalue96S3, error96S3, CLU96S3, CLL96S3, wd96S3)


# Calculating Confidence for 99 %
####################################################
CL99S3 <- 0.99

# Calculation of Z Value for 99% CI
zvalue99S3 <- qnorm((1 + CL99S3) / 2)

# Calculation of Error for 99% CI
error99S3 <- zvalue99S3 * sqrt(((p_successS3 * p_failureS3) / nS3))

# Calculation of Upper and Lower Limit for 99% CI
CLL99S3 <- p_successS3 - error99S3
CLU99S3 <- p_successS3 + error99S3

# Calculation of C.I. Width for 99%
wd99S3 <- CLU99S3 - CLL99S3

# Vector Creation for 99% C.I.
vectorCI99S3 <- c(zvalue99S3, error99S3, CLU99S3, CLL99S3, wd99S3)


# Matrix Creation for compiling all the Confidence Interval Vectors
#########################
S3_Table <- matrix(c(vectorCI90S3, vectorCI96S3, vectorCI99S3),
  nrow = 3,
  byrow = TRUE
)


## Rounding Digits in Matrix
S3_Table_Rounded <- round(S3_Table, digits = 2)


## Naming Columns and Row Names in Table
colnames(S3_Table_Rounded) <- c("Z Value", "Mar of Error", "Upper C.L.", "Lower C.L.", "Width of C.I.")

rownames(S3_Table_Rounded) <- c("CI 90%", "CI 96%", "CI 99%")

# Table Creation of of Matrix Using Knitr Package
S3_Table_Rounded %>%
  knitr::kable(
    caption = "Table 7: Results of Sample 3 With Different Confidence Interval",
    format = "html"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kable_classic(
    full_width = F,
    html_font = "Times New Roman"
  )
```


<BR>
<FONT SIZE = 4.5>

***Observations***

</FONT>


<P>
<BR> <B>
<FONT SIZE = 4.75, COLOR ="#f94c56">
2.7. Sample 4 : Confidence Interval for Sample Variance
</FONT>
</BR></B>


```{r message=FALSE, warning=FALSE}


# Sample 4 : Basic Descriptive Statistics
nS4 <- 150
VarS4 <- var(S4)
StdS4 <- sd(S4)

# Calculating the Variance for Population
VarPop <- var(SData$Population)

# Inline R Code to Print Population and Sample Variance
paste("The variance of Sample 4 is", round(VarS4, 2))
paste("The variance of Population is", round(VarPop, 2))


# Creating Object for 90% Confidence Interval for Sample 4 Variance
########################################
CL90S4 <- 0.90
dfS4 <- nS4 - 1


# Calculating Alpha Value
alp90S4 <- (1 - CL90S4)

# Calculating Z value for Variance

Z90LS4 <- qchisq(p = alp90S4, df = dfS4, lower.tail = TRUE)

Z90RS4 <- qchisq(p = alp90S4, df = dfS4, lower.tail = FALSE)




# Creating Objects for Confidence Interval: Upper and Lower Limit

CLL90S4 <- ((dfS4) * (StdS4^2)) / Z90LS4

CLU90S4 <- ((dfS4) * (StdS4^2)) / Z90RS4

# Calculation of C.I. Width for 90
wd90S4 <- CLL90S4 - CLU90S4

# Inline R Code to present all the value calculated for 90% CI
paste("1.1. The -ve Z Value of Sample 4 with 90% Confidence Interval is", round(Z90LS4, 2))

paste("1.2. The +ve Z Value of Sample 4 With 90% Confidence Interval is", round(Z90RS4, 2))

paste("1.3. The Confidence Upper Limit for Sample 4 with 90% Confidence Interval is", round(CLU90S4, 2))

paste("1.4. The Confidence Lower Limit for Sample 4 with 90% Confidence Interval is", round(CLL90S4, 2))

paste("1.5. The Confidence Interval Width for Sample 4 with 90% Confidence Interval is", round(wd90S4, 2))


# Creating Object for 96% Confidence Interval for Sample 4 Variance
########################################
CL96S4 <- 0.96
dfS4 <- nS4 - 1


# Calculating Alpha Value
alp96S4 <- (1 - CL96S4)

# Calculating Z value for Variance

Z96LS4 <- qchisq(p = alp96S4, df = dfS4, lower.tail = TRUE)

Z96RS4 <- qchisq(p = alp96S4, df = dfS4, lower.tail = FALSE)




# Creating Objects for Confidence Interval: Upper and Lower Limit

CLL96S4 <- ((dfS4) * (StdS4^2)) / Z96LS4

CLU96S4 <- ((dfS4) * (StdS4^2)) / Z96RS4

# Calculation of C.I. Width for 96
wd96S4 <- CLL96S4 - CLU96S4

# Inline R Code to present all the value calculated for 96% CI
paste("2.1. The -ve Z Value of Sample 4 with 96% Confidence Interval is", round(Z96LS4, 2))

paste("2.2. The +ve Z Value of Sample 4 With 96% Confidence Interval is", round(Z96RS4, 2))

paste("2.3. The Confidence Upper Limit for Sample 4 with 96% Confidence Interval is", round(CLU96S4, 2))

paste("2.4. The Confidence Lower Limit for Sample 4 with 96% Confidence Interval is", round(CLL96S4, 2))

paste("2.5. The Confidence Interval Width for Sample 4 with 96% Confidence Interval is", round(wd96S4, 2))

# Creating Object for 99% Confidence Interval for Sample 4 Variance
########################################
CL99S4 <- 0.99
dfS4 <- nS4 - 1


# Calculating Alpha Value
alp99S4 <- (1 - CL99S4)

# Calculating Z value for Variance

Z99LS4 <- qchisq(p = alp99S4, df = dfS4, lower.tail = TRUE)

Z99RS4 <- qchisq(p = alp99S4, df = dfS4, lower.tail = FALSE)




# Creating Objects for Confidence Interval: Upper and Lower Limit

CLL99S4 <- ((dfS4) * (StdS4^2)) / Z99LS4

CLU99S4 <- ((dfS4) * (StdS4^2)) / Z99RS4

# Calculation of C.I. Width for 99
wd99S4 <- CLL99S4 - CLU99S4

# Inline R Code to present all the value calculated for 99% CI
paste("3.1. The -ve Z Value of Sample 4 with 99% Confidence Interval is", round(Z99LS4, 2))

paste("3.2. The +ve Z Value of Sample 4 With 99% Confidence Interval is", round(Z99RS4, 2))

paste("3.3. The Confidence Upper Limit for Sample 4 with 99% Confidence Interval is", round(CLU99S4, 2))

paste("3.4. The Confidence Lower Limit for Sample 4 with 99% Confidence Interval is", round(CLL99S4, 2))

paste("3.5. The Confidence Interval Width for Sample 4 with 99% Confidence Interval is", round(wd99S4, 2))
```

<BR>
<FONT SIZE = 4.5>

***Observations***

</FONT>


<P>
<BR> <B>
<FONT SIZE = 4.75, COLOR ="#f94c56">
2.8. Sample 5 : Hypothesis Testing Using Using z-Test
</FONT>
</BR></B>

```{r message=FALSE, warning=FALSE}

# Creating Objects for Sample 5
nS5 <- 200
MeanS5 <- mean(S5)

# Creating Objects for Population
MeanPop <- mean(SData$Population)
StdPop <- sd(SData$Population)

CL95S5 <- 0.95

# Calculating Alpha Value
alp95S5 <- (1 - CL95S5)

# Z-Test
ZtestS5 <- (MeanS5 - MeanPop) / (StdPop / sqrt(nS5))

ZtestS5

# Calculating Right Critical value as Z-test is positive
Z95LS5 <- qnorm(alp95S5 / 2)
Z95RS5 <- qnorm((alp95S5 / 2) + CL95S5)

# Checking if Z-Test Value is greater than right critical value
ZtestS5 > Z95RS5

vectorS5 <- c(CL95S5, Z95LS5, Z95RS5, ZtestS5)



HypothesisMatrix <- matrix(
  data = c(vectorS5),
  nrow = 1,
  byrow = TRUE
)

## Rounding Digits in Matrix
HypothesisMatrix_Rounded <- round(HypothesisMatrix, digits = 2)

colnames(HypothesisMatrix_Rounded) <- c("Confidence Interval", "Left Critical Value", "Right Critical Value", "Z-test Value")

# Present the table using kableExta Package
knitr::kable(HypothesisMatrix_Rounded,
  caption = "Table 8: Sample 5: Result of Hypothesis Testing (Z-Test) ",
  format = "html",
  table.attr = "style=width: 40%", align = "lccrr",
  font_size = 8
) %>%
  kable_styling(bootstrap_options = c(
    "striped", "hover",
    "condensed", "responsive"
  )) %>%
  kable_classic(
    full_width = F,
    html_font = "Times New Roman"
  )
```

<BR>
<FONT SIZE = 4.5>

***Observations***

</FONT>


<P>
<BR> <B>
<FONT SIZE = 4.75, COLOR ="#f94c56">
2.9. Sample 5 : Hypothesis Testing Using P-Value Method
</FONT>
</BR></B>


```{r message=FALSE, warning=FALSE}

# Calculating P Value from Z-test
pvalueS5 <- round(1 - pnorm(ZtestS5), digits = 3)


# Checking if the p-value is less than alpha value on right-tail
pvalueS5 < alp95S5 / 2
```

<BR>
<FONT SIZE = 4.5>

***Observations***

</FONT>


<P>
<BR> <B>
<FONT SIZE = 4.75, COLOR ="#f94c56">
2.10. Sample 6 : Hypothesis Testing Using t-Test
</FONT>
</BR></B>


```{r message=FALSE, warning=FALSE}

# Creating Objects for Sample 5
nS6 <- 29
MeanS6 <- mean(S6)
alphaS6 <- 0.01

# Creating Objects for Population
MeanPop <- mean(SData$Population)
StdPop <- sd(SData$Population)


# calculating Degree of freedom
dfS6 <- nS6 - 1

# Calculating t-value
tvalueS6 <- qt(1 - alphaS6, dfS6)

# Performing the right tailed test based on hypothesis
ttestS6 <- (MeanS6 - MeanPop) / (StdPop / sqrt(nS6))

ttestS6 > tvalueS6

# Calculating P Value from t-test
pvalueS6 <- round(1 - pnorm(ttestS6), digits = 3)


pvalueS6 < alphaS6

vectorS6 <- c(alphaS6, tvalueS6, ttestS6, pvalueS6)

HypothesisMatrix6 <- matrix(data = c(vectorS6), nrow = 1, byrow = TRUE)

## Rounding Digits in Matrix
HypothesisMatrix6_Rounded <- round(HypothesisMatrix6, digits = 2)

colnames(HypothesisMatrix6_Rounded) <- c("Alpha Value", "T-Value", "Test Value", "P-Value")

# Present the table using kableExta Package
knitr::kable(HypothesisMatrix6_Rounded,
  caption = "Table 9: Sample 6: Result of Hypothesis Testing (t-Test) ",
  format = "html",
  table.attr = "style=width: 40%", align = "lccrr"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

<BR>
<FONT SIZE = 4.5>

***Observations***

</FONT>

<P>
<BR> <B>
<FONT SIZE = 5, COLOR ="#f70d1a">
3. CONCLUSIONS
</FONT>
</BR></B>

- 





<P>
<BR> <B>
<FONT SIZE = 5, COLOR ="#f70d1a">
4. BIBLIOGRAPHY
</FONT>
</BR></B>

<P>
<BR> <B>
<FONT SIZE = 5, COLOR ="#f70d1a">
5. APPENDIX
</FONT>
</BR></B>

An R Markdown file has been attached to this report. The name of the
file is Maheswar_M3Project.Rmd